---
output: github_document
---

[![stability-experimental](https://img.shields.io/badge/stability-experimental-orange.svg)](https://github.com/joethorley/stability-badges#experimental)
[![Travis-CI Build Status](https://travis-ci.org/cboettig/nexld.svg?branch=master)](https://travis-ci.org/cboettig/nexld) 

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA,
  fig.path = "README-"
)
```

# nexld

The goal of nexld is to serialize NeXML xml data into JSON-LD.

## Installation

You can install nexld from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("cboettig/nexld")
```

## Heads up!

Package is in purely exploratory stage at this time, function api likely to change.


## Example

Let's convert a simple NeXML file into JSON-LD:

```{r example}
library(nexld)
ex <- system.file("extdata/example.xml", package = "nexld")

xml_to_json(ex)
```

Obviously this is just a preliminary mapping to start discussion.  The convention is relatively self-explanatory.  Repeated nodes (`edge`, `otu` etc) are replaced by a list of such nodes.  


Properties prefixed with `@` symbol, such as `@id`, `@type`, `@context`, are special in JSON-LD.  The `@context` plays the role of the namespace and schema (e.g. also defines expected data types); in this example I haven't defined the context.  Most of the basic JSON-LD operations can be seen as manipulations based on the `@context`.  Pasting the above JSON-LD into <https://json-ld.org/playground> is a good way to explore these functions.  The same methods are also available in R through the `jsonld` R package.  



### Questions / issues

1.  For the moment nodes are typed according to their xml names; though other typing schemes could be used (might make more sense to use the classes directly from the XSD schema?  Schema.org convention always begins types with a capital letter instead.)

2.  It would no doubt make more sense to replace meta tags with their intended expansion; e.g. instead of nodes having a property `meta` which takes `"@type": "meta"` nodes, meta nodes should be expanded, e.g.: 

```
            {
              "@type": "node",
              "@id": "tree2n4",
              "about": "#tree2n4",
              "label": "n4",
              "meta": {
                "@id": "tree2dict1",
                "property": "cdao:has_tag",
                "content": "true",
                "@type": "nex:LiteralMeta",
                "datatype": "xsd:boolean"
              }
```

Would simply become

```
            {
              "@type": "node",
              "@id": "tree2n4",
              "about": "#tree2n4",
              "label": "n4",
              "cdao:has_tag": true
              }
```

Note that the data type would be defined in the context file for `cdao:has_tag`. Similarly, a ResourceMeta would probably just become a URI-valued property; e.g. 

```
  <meta xsi:type="ResourceMeta" href="http://creativecommons.org/publicdomain/zero/1.0/" rel="cc:license"/>
```

to:

```
"cc:license": "http://creativecommons.org/publicdomain/zero/1.0/"
```


3. The `about` elements should not be treated just as ordinary properties, since they define the subject of the triple.  While this was necessary in the XML to make the relationship explicit (rather than implicit in the graph structure), in JSON-LD this is inferred directly from nesting.   e.g. in the following block:

```
"tree": {
  "@id": "tree1",
  "node": [{
    "@id": "tree1n1",
    "about": "#tree1",
    "label": "n1"
  }]
}
```

The use of `about` is redundant; (or worse, it would be conflicting for that node to be `about` anything other than its parent).  JSON-LD includes a notion of flattening, where all nodes are unnested and refered to only by `@id`, e.g. flattening the above creates the semantically equivalent:

```
"tree": {
  "@id": "tree1",
  "node": [ "tree1n1"]
},
{
  "@id": "tree1n1"
  "label": "n1"
}
```

Note that `@id` elements are not required, if not available JSON-LD will assign relative ids using the `_:b0`, `_:b1` ... id scheme.  This suggests that we should: (a) make sure the parent element `id` matches that given in the `about` node, and if not, then relocate the element, and then (b) simply drop the `about` attributes.  

4. We would probably want to write out the actual context file for all the NeXML types.  This isn't essential; but provides declarations of the data type.    

5.  There are probably other issues that can arise that I haven't thought about.

6. So far this is just a one-way conversion to establish the convention, though I think a roundtrip mapping should be possible.  One of the challenges in the reverse direction of the meta expansion, since it requires recognizing that the property isn't in the NeXML namespace (or maybe that's easy since we have the namespaces)
